\documentclass{article}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{semantic-markup}
\usepackage{musicography}
\usepackage{fancyvrb}
\frenchspacing

\newenvironment{codetable}
{\begin{quote}\begin{tabular}{lll}}
{\end{tabular}\end{quote}}

\usepackage[
    pdftitle={The musicography Package: Symbols for Music Writing with
    pdflatex},
    pdfauthor={Andrew A. Cashner},
    pdfsubject={LaTeX package},
    pdfkeywords={LaTeX, symbols, music, musicology, humanities}
]{hyperref}

\title{The \code{musicography} Package: Symbols for Music Writing with
\code{pdflatex}}
\author{Andrew A. Cashner\thanks{%
    \href{mailto:andrewacashner@gmail.com}
    {\nolinkurl{andrewacashner@gmail.com}}%
    }%
}

\begin{document}
\maketitle

Font packages for \code{pdflatex} only provided a limited range of musical
symbols. 
The \code{lilyglyphs} package uses Lilypond's fonts, but requires
\code{lualatex}. 
This package makes available the most commonly used symbols in writing about
music in a way that can be used with \code{pdflatex} and looks consistent and
attractive.
It includes accidentals, meters, and notes of different rhythmic values.

This package builds on the approach used in the \code{harmony} package, where
the symbols are taken from the MusiXTeX fonts.
But it provides a larger range of symbols and a more flexible, user-friendly
interface written using \code{xparse}.

\tableofcontents

\section{Package Options}

To use the package, write \verb|\usepackage{musicography}| in your preamble.
If you are also using this author's \code{semantic-markup} package, load
\code{semantic-markup} first, since \code{musicography} modifies the commands
for accidentals in the other package.

The \code{bigger} option provides larger font sizes that match better with
certain fonts.

The \verb|\musNumFont| command allows you to change the font of the numerals
used.
For example, if using the \code{ebgaramond} font package, it looks better if you
do this: 
\begin{quote}
    \verb|\renewcommand{\musNumFont}[1]{\liningnums{#1}}|.
\end{quote}

\section{Symbols and Commands}

\subsection{Accidentals}

\begin{codetable}
    Flat & \verb|\musFlat| or \verb|\fl| & \musFlat\\
    Sharp & \verb|\musSharp| or \verb|\sh| & \musSharp\\
    Natural & \verb|\musNatural| or \verb|\na| & \musNatural\\
    Double Flat & \verb|\musDoubleFlat| & \musDoubleFlat\\
    Double Sharp & \verb|\musDoubleSharp| & \musDoubleSharp\\
\end{codetable}

\subsection{Notes of Different Rhythmic Values}

Commands are available using modern (United States) note names; in several cases
there are also aliases for older note names.

\begin{codetable}
    Whole note (semibreve) & \verb|\musWhole| or \verb|\musSemibreve| &
    \musWhole\\
    Half note (minim) & \verb|\musHalf| or \verb|\musMinim| & \musHalf\\
    Quarter note (semiminim) & \verb|\musQuarter| or \verb|\musSeminimin| &
    \musQuarter\\
    Eighth note (corchea) & \verb|\musEighth| or \verb|\musCorchea| &
    \musEighth\\
    Sixteenth note & \verb|\musSixteenth| & \musSixteenth\\
\end{codetable}

A dot may be added to any of the above by adding \code{Dotted} to the end of the
command. For example:

\begin{codetable}
    Dotted whole note & \verb|\musWholeDotted| & \musWholeDotted\\
    Dotted quarter note & \verb|\musQuarterDotted| & \musQuarterDotted\\
\end{codetable}

\subsection{Meter Signatures}

\begin{codetable}
    Common duple & \verb|\meterC| & \meterC\\
    \term{Alla breve} & \verb|\meterCutC| & \meterCutC\\
    Ternary (16th--18th cent.) & \verb|\meterCThree| & \meterCThree\\
    Ternary with $3:2$ proportion & \verb|\meterCThreeTwo| &
    \meterCThreeTwo\\
    Spanish 17th-cent. ternary & \verb|\meterCZ| & \meterCZ\\
\end{codetable}

\subsection{Customization}

It would be a simple matter of using \verb|\newcommand| or \verb|\let| to create
aliases for these commands, say, for British usage (such as \verb|\quaver|).

\LaTeX{} programmers may wish to use the package's internal commands directly 
(such as \verb|\musSymbol| and \verb|\musAccidental|) to
access more symbols from the fonts or to fine-tune their appearance (see the
code listing below).

\section{Code}

\VerbatimInput{musicography.sty}

\end{document}
